\bf DOGE's Downsizing, Can AI Read the Reddit Room?

### 2025-04-08 code/ml_llm/R/assign_3.tex
adding paper

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={DOGE Days on Reddit: Decoding Public Sentiment in a Federal Shakeup},
  pdfauthor={Kevin Linares; Felix Baez-Santiago; Aria Lu; Gloria Zhou},
  pdfkeywords={Reddit, Federal Government, DOGE},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\date{April 7, 2025}
\title{\bf DOGE Days on Reddit: Decoding Public Sentiment in a Federal
Shakeup}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
This study investigates public sentiment on Reddit towards the
Department of Government Efficiency's (DOGE) federal workforce
reduction. Analyzing 400 labeled Reddit comments, we employed supervised
learning (KNN, Random Forest) and unsupervised Large Language Models
(LLMs: Llama 3.2 3B, deepseek-r1:8b) to classify comment stance (favor,
neutral, oppose). While LLMs showed higher precision for opposing
comments, all models struggled with neutral and favoring stances.
Agreement between the two LLMs was fair (Cohen's Kappa = 0.344),
indicating divergent interpretations. Results suggest that accurately
decoding nuanced public sentiment in this context remains challenging
for both traditional and large language models.
\end{abstract}

@@

Established in January 2025, the Department of Government Efficiency
(DOGE) has significantly reduced the federal workforce by over 280,000
employees, fulfilling a key campaign promise of the new administration.
This has caused widespread concern and anxiety among federal workers
about job security and mental well-being. Questions have also been
raised regarding DOGE's adherence to internal policies. To understand
the impact of DOGE's actions on federal worker perceptions of job
security, we collected 400 reddit comments on topics related to DOGE's
tactics in reducing the federal workforce. Additionally, a team of
researchers labeled each comment as whether the author favored, opposed,
or had a neutral stance concerning how DOGE was implementing these
strategies. In the current study we employ two supervise learning model
to predict, given a set of features, comments on stance. We also employ
an unsupervised large language model (LLM) to detect the stance of these
reddit comments and report model findings.

@@

\emph{Data}. We collected Reddit comments from subreddits related to the
topic of interest in early March of 2024. This resulted in 12,553
comments which we preprocessed by removing weblinks, replace special
characters (i.e., replaced ``@'' with ``at''), replaced numeric values
with their spelling, and confirmed the absence of duplicate comments. We
then randomly selected 400 comments (without replacement) and assigned
them to four graduate students for coding, categorizing each comment as
favoring, opposing, or neutral towards DOGE's approach to federal
workforce reductions. Table 1 presents the breakdown in percent of our
the comments we labelled and later use to train and evaluate our machine
learning models. A review of the comments revealed more negative
opinionated statements and reactions.

@@

We built two supervised machine learning models, K-nearest neighbor and
random forest, to classify Reddit comments concerning stance on the
current reduction in federal workforce into favor, oppose, or neutral.
For model predictors we used the Reddit score, up-votes, and down-votes
each comment received at the time of the collection. Additionally, we
used the keyword that was used to scrape these data along with the
subreddit where the comment was posted. DISCUSS ADDITIONAL FEATURE
ENGINERRING SUCH AS TEXT EDITING THAT WENT INTO THE MODEL. For the KNN
model, we set the number of neighbors hyperparameter to thhree, and for
the random forest, mtry was kept at a constant two as we did not have
many features in these models. We took an 80/20 split for our
training-testing sets that we used to train each model and evaluate.

In addition to our supervised machine learning models, we also applied
to our data two LLMs, llama 3.2. 3B and deepseek-r1:8b. The llama model,
developed by Meta AI released in February 2023, was designed for
applications requiring low-latency inferencing with limited
computational resources, which means that we were able to also deploy
these models on a local machine with a GPU in addition to using the
\href{https://its.umich.edu/advanced-research-computing/}{Great Lakes
HCP}. This model also excels at text classification which is how we used
it to classify the stance for each comment. Deepseek, developed by Liang
Wenfeng in July 2023 and based on the llama architecture, yet boosts
efficiency improvements that appear to provide advances in reasoning,
logic, while excelling at natural language processing. We choose both of
these LLMs as candidates to test against our supervised approaches
because of their ability to apply locally and reputable performance on
zero-shot classification tasks. Both of these models received the same
prompt and tasked which were developed by the research team to reflect
the utility of classifying stance within our topic of interest. Both
LLMs resulted in not being able to classify five comments each while
they only had one in common with the likely reason being that these
comments did not contain enough useful text for a stance to be derive.

Prompt: ``Is this comment in `favor', `neutral', or `oppose' the
reduction in federal workforce? Provide one word answer only!

Task: ``You have assumed the role of a stakeholder that is presented
with a reddit comment from a likely federal worker related to the
current policies on reducing the federal workforce. Please determine the
author's stance on this topic, and only provide the answer.''

To evaluate model performance we used recall to determine how many of
actual positive cases did the model correctly identify. High recall
indicates that the model can correctly identify positive instances but
is also optimistic and may incorrectly label cases as positive. The
second performance metric we selected is precision, of all the items the
model labels as positive how many are true positive, and high precision
indicates that when the model predicts a case to be positive it is
likely to be correct.

@@

Our machine learning models reveal a mix of performance results in
classifying comments into one of three stance categories. Overall, the
random forest model reduces the test misclassification error against
KNN, yet not by much. The stance oppose tended to have the highest
recall of the three with the random forest model having higher scores:
Of all the oppose cases, the random forest model found 74 percent and
the KNN 64 percent. For the neutral stance, the random forest captured 5
percent and KNN 21 percent. While for the favor stance, the random
forest identified 26 percent and KNN 16 percent (see Figure 1 for these
results). Neither model excelled in how many of the stance categories
identified were right as indicated by the precision scores. KNN slightly
outperformed random forest on the oppose stance having correctly
classified 56 percent vs random forest 51 percent or true oppose cases.
For neutral the performance is worse with random forest correctly
identified 50 percent and KNN 21 percent, and same for favor with random
forest correctly identifying 29 percent and KNN 23 percent.

@@

We observe higher precision scores for the oppose stance when using the
LLMs to classify stance positions. Of all the identified oppose stance
cases these models identified, the percent of true oppose cases was 84
percent for llama and 69 percent for deepseek. Similarly, of all the
classified neutral stance classified by these models, the percent of
true neutral cases was 21 percent for llama and 42 percent for deep
seek. Finalkly, of all the identified favor stance, the percent of true
favor cases was 4 percent for llama and 10 percent for deepseek (see
Figure 2 for these results). Recall scores were worse for the oppose
stance, and of all the oppose cases the llama model captured 56 percent
and deepseek 60 percent. For the neutral stance, llama captured 25
percent and deekseek 27 percent. Finally, for the favor stance, llama
captured 29 percent and deekseek 32 percent.

\hfill\break

@@

Llama and DeepSeek demonstrated only fair agreement in stance
classification (Cohen's Kappa = 0.344). The contingency table, Table 2
below, reveals highest agreement for ``oppose'' (0.58), indicating some
consistency in identifying strong negative sentiment. However, agreement
was substantially lower for ``neutral'' (0.11) and ``favor'' (0.02),
highlighting divergent interpretations of more nuanced language.
Off-diagonal values further illustrate these discrepancies, suggesting
fundamental differences in how the models process ambiguous cues and
establish classification boundaries, particularly for less extreme
stances.

\begin{longtable}[]{@{}lrrr@{}}
\caption{Aggreement between LLMs, Cohens Kappa, 0.34}\tabularnewline
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
favor & 6 & 7 & 1 \\
neutral & 3 & 42 & 14 \\
oppose & 29 & 61 & 228 \\
\end{longtable}

@@

We explored public sentimenton Reddit regarding the DOGE federal
workforce reduction using supervised learning (KNN, Random Forest) and
unsupervised LLMs (Llama 3.2 3B, deepseek-r1:8b) for stance
classification. LLMs showed higher precision for ``oppose'' comments,
suggesting better identification of negative sentiment, but lower
recall, indicating missed opposing comments. Supervised models performed
consistently but generally poorly across all stances, and likely due to
the limited feature set (Reddit metadata) as well as the small sample of
comments.

The consistent difficulty in classifying ``neutral'' and ``favor''
stances across all models likely stems from class imbalance (more
``oppose'' comments), the inherent ambiguity of neutral language, and
potentially more nuanced expressions of favor. The zero-shot approach
for LLMs might also limit their performance on this specific domain. The
few unclassifiable comments highlight the inherent challenges in stance
detection within informal online text.

We conclude that while LLMs showed promise for identifying strong
negative sentiment, accurately classifying the full spectrum of public
opinion on Reddit regarding DOGE's actions remains challenging. Future
research should focus on incorporating richer text-based features for
supervised models, addressing class imbalance, and potentially
fine-tuning LLMs to better capture the nuances of online sentiment and
stance.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={DOGE's Downsizing, Can AI Read the Reddit Room?},
  pdfauthor={Kevin Linares; Felix Baez-Santiago; Aria Lu; Gloria Zhou},
  pdfkeywords={Reddit, Federal Government, DOGE},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\date{April 8, 2025}
\title{\bf DOGE's Downsizing, Can AI Read the Reddit Room?}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
We investigate public sentiment on Reddit regarding the Department of
Government Efficiency's federal workforce reduction by classifying 400
labeled comments (28\% favor, 18\% neutral, 54\% oppose) using
supervised and unsupervised Large Language Models. Supervised models
showed moderate success, particularly with ``oppose'' comments, but
struggled with ``favor'' and ``neutral'' stances. Similarly, LLMs best
identified ``oppose'' sentiment but exhibited low precision for
``favor'' and ``neutral.'' These findings highlight the challenges of
accurately gauging nuanced public opinion on government policy changes
using automated methods on social media data.
\end{abstract}

@@

The newly formed Department of Government Efficiency (DOGE) has reduced
the federal workforce by almost 280,000 employees, a central pledge of
the current administration. This action has generated significant
apprehension among federal workers in regards to mental health and job
security. To understand the impact of DOGE's actions on federal worker
perceptions of job security, we labeled 400 Reddit comments on topics
related to the current reduction in \emph{federal} workforce by DOGE as
whether the author favored, opposed, or had a neutral stance. We use
these labels to build supervised learning models to predict stance.
Additionally, we employ unsupervised large language models (LLM) to
detect the stance of these Reddit comments from text, to further explore
appropriate models for this current topic.

@@

\emph{Data}. We collected Reddit comments from subreddits related to the
topic of interest in early March of 2024. This resulted in 12,553
comments which we preprocessed by removing web-URLs, replace special
characters (i.e., replaced ``@'' with ``at''), replaced numeric values
with their spelling, and confirmed the absence of duplicate comments. We
then randomly selected 400 comments (without replacement) and assigned
them to four graduate students for coding, categorizing each comment as
favoring, opposing, or neutral towards DOGE's approach to federal
workforce reductions. Table 1 presents the breakdown in percent of our
the comments we labelled and later use to train and evaluate our machine
learning models. A review of the comments revealed more negative
opinionated statements and reactions.

@@

We built two supervised machine learning models--K-nearest neighbor and
random forest--to classify Reddit comments concerning stance on the
current reduction in federal workforce into favor, oppose, or neutral.
For predictors we used the Reddit score, up-votes, and down-votes each
comment received at the time of the collection. Additionally, we used
the keyword that was used to scrape these data along with the subreddit
where the comment was posted. \textbf{DISCUSS ADDITIONAL FEATURE
ENGINERRING SUCH AS TEXT EDITING THAT WENT INTO THE MODEL.} For the KNN
model, we set the number of neighbors hyperparameter to three, and for
the random forest, mtry was kept at a constant two as we did not have
many features in these models. We took an 80/20 split for our
training-testing sets that we used to train each model and evaluate.

In addition to our supervised machine learning models, we also applied
to our comments two LLMs, gemma 3.12 and llama 3.2. 3B. Gemma, was
developed by Google and is based on Gemini, is able to generate text
output using instruction-tuned prompts from the user, such as providing
answers on a user specified task reviewing text. We deployed gemma on a
higher performance computing cluster
(\href{https://its.umich.edu/advanced-research-computing/}{Great Lakes
HCP}) and processed text from Reddit comments on a GPU. We were also
interested in deploying other lightweight LLMs locally on a machine with
a dedicated-GPU and found llama, developed by Meta AI, to be an
excellent candidate. Llama excels at text classification which is how we
used it to classify the stance for each comment text given a user
specified prompt. We choose both of these LLMs as candidates to test
against our supervised approaches because of their ability to apply
locally and reputable performance on zero-shot classification tasks.
Both of these models received the same prompt and tasked which were
developed by the research team to reflect the utility of classifying
stance within our topic of interest.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2222}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7778}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Prompt}: & ``Is this comment in `favor', `neutral', or `oppose'
the reduction in federal workforce? Provide one word answer only! \\
\textbf{Task}: & ``You have assumed the role of a stakeholder that is
presented with a reddit comment from a likely federal worker related to
the current policies on reducing the federal workforce. Please determine
the author's stance on this topic, and only provide the answer.'' \\
\end{longtable}

We evaluate model performance using the F1-score which is a harmonic
mean balance between precision and recall. This metric considers both
the ability to correctly identify positive instances, known as recall,
and the ability to avoid incorrectly labeling negative instances as
positive (i.e., precision). A high F1-score indicates a model with well
balanced recall and precision that both minimize false negativesand
false positives. The second performance metric we used is recall, as
this determines how many of the actual positive cases the model can
correctly identify. Recall can be prone to incorrectly labeling negative
cases as positive, suggesting more false positives, which is why we
intereggoate on two performance metrics in this study.

@@

Our machine learning models reveal a mix of performance results in
classifying the stance of Reddit comment into favor, neutral, oppose.
Both models exhibit low recall and F1-scores for the ``favor'' stance
(see Figure 1), yet the random forest model (recall .26: f1-score .28)
slightly outperforms KNN (recall .16: f1-score .19). The ``neutral''
class presented a significant challenge for the random forest with very
low scores (recall .05: f1-score .10), while KNN exhibited slightly
better performance (recall .21: f1-score .21). Both models showed the
strongest performance in identifying comments expressing opposition. The
random forest correctly identified 74 percent of all the actual Reddit
comments expressing opposition compared to KNN at 64 percent, meaning
that KNN missed a larger proportion of true ``oppose'' comments. The
f1-scores were similar for both models at 60 percent, this implies that
KNN may have a slightly higher precision score given that it had a lower
recall score than the random forest model. This implies that when KNN
predicts comments for the ``oppose'' class, it is as likely as the
random forest to be correct.

@@

The LLMs, showed poor performance in classifying the ``favor'' stance,
with very low f1-scores (llama .07: gemma .02) despite moderate recall
(llama .29: gemma .33), indicating low precision (see Figure 2). For the
``neutral'' class, both models showed modest results, with llama
achieving a slightly better F1-score (.23) compared to gemma (0.18),
although gemma had a higher recall (.50 vs .25), again suggesting lower
precision for gemma. Similar to the supervised models, both LLMs
performed best in classifying the ``oppose'' stance, with relatively
high and similar F1-scores (llama .67: gemma .70) and comparable recall
(llama .56: gemma .56), suggesting a better balance between precision
and recall for this category. Both models correctly identify 67 to 70
percent of all the actual Reddit comments expressing opposition.
Overall, both LLMs struggled with the ``favor'' and ``neutral'' classes
but demonstrated a stronger ability to identify opposing comments.\\

@@

Gemma and llama demonstrated only fair agreement in stance
classification (Cohen's Kappa = .24). The contingency table, Table 2
below, reveals highest agreement for ``oppose'' (.80), indicating some
consistency in identifying a strong negative stance. However, agreement
was substantially lower for ``neutral'' (.03) and ``favor'' (.01),
highlighting divergent interpretations of more nuanced language.
Off-diagonal values further illustrate these discrepancies, suggesting
fundamental differences in how the models process ambiguous cues and
establish classification boundaries, particularly for less extreme
stances.

\begin{longtable}[]{@{}lrrr@{}}
\caption{Aggreement between LLMs, Cohens Kappa, 0.24}\tabularnewline
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
favor & 3 & 1 & 10 \\
neutral & 0 & 10 & 49 \\
oppose & 0 & 5 & 317 \\
\end{longtable}

@@

We explored public sentiment on Reddit regarding the DOGE federal
workforce reduction using supervised learning (KNN, Random Forest) and
unsupervised LLMs ( gemma 3.12b and Llama 3.2 3B) for stance
classification. Our supervised learning models achieved moderate success
in this classification task, with the strongest performance in
identifying opposing viewpoints, which were also the most prevalent in
our labeled data. Moreover, both models struggled with the ``favor'' and
``neutral'' stances, suggesting limitations in our initial features to
capture these nuances. Our unsupervised learning models also showed the
best ability to classify ``oppose'' comments, yet exhibited significant
challenges with the ``favor'' and ``neutral'' categories, particularly
demonstrating low precision.

This study had limitations such as the small size of our labelled
dataset of 400 Reddit, imbalance in the stance categories, and few
features in the supervised approach. The zero-shot capabilities of the
LLMs are suitable for topic exploration but may require fine-tuning for
optimal performance on this specific domain, particularly with nuance
language used in ``favor'' comments. Future research should continue to
understand federal worker perceptions and improve stance classification
by expanding labeled dataset and incorporate more contextual text rather
than single comment text. Fine tuning LLMs on a larger corpus with
relevant comments could render more accuracy in the classification task.
